Stack Safety as a Security Property

### 2020-07-08 main.tex
pass on 5.1 and 5.2; releasing lock on them; keeping lock on 5.3

In this section we will describe stack safety as a formal property. We
begin with our model of \leo{compiler-provided?} knowledge about the
program structure, and then dive into the formal definitions of local state
encapsulation and well-bracketed control flow that together constitute
stack safety.

@@

\paragraph*{Return Map}
Similarly, a {\em return map} identifies addresses that complete a
return sequence. Formally, it is simply a set of components; each one
is the final jump corresponding to a function:
%
\[\retmap \in \RETMAPS \subseteq \WORDS\]
%
In our running example, the return map identifies the $\jalr$
instructions 36 and 47:
\[ \retmap = \{36, 47\} \]

@@

We separately identify, for any call state, what it means for a state
to have correctly returned from a call. We consider the call to have
returned the first time control returns to the instruction following
the \jal with the stack restored. Formally this is a relation
between the states:

\[\begin{split}
    \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) =
    \nextPC(\mach_c(\PCname)) \land \\ & \mach_r(\SP) = \mach_c(\SP)
\end{split}\]

There are two such return points in our running example, both of which
are depicted in the figure: the state before the execution of
instruction 27 corresponding to the return from {\tt h} and the state
before instruction 8 corresponding to the return from {\tt g}.

@@

In the rest of this section we will implictly parameterize our
definitions by a function map $\codemap$, a call map $\callmap$, a return
map $\retmap$ and an entry map $\entmap$.

@@

When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) and
\(\mach_c \stepstoobs{\obs} \mach_e\) for some $n$, $\obs$, and
$\mach_e$, we can construct a contour based on the bounds of the
callee's stack frame.
%
The private state of the callee is the entire area of the memory from
stack base to stack pointer, with the exception of the $n$ locations
corresponding to the arguments and result locations of the
call. Private components are marked high confidentiality and high
integrity: the callee should be able to neither read, nor write them.
%
Everything above the stack pointer at the point of the call is
``uninitialized'' memory that can be written into but should not be
read. To protect the callee itself from such reads, these components
are marked high confidentiality and low integrity.
%
The same is true for the designated output location: it can be
freely written into but should not be read.
%
Special care is needed for the code memory corresponding to the
callee: they are marked low confidentiality if they belong to the
callee and high confidentiality otherwise; regardless they are always
high integrity.
%
Finally, registers and the frame of arguments to the callee are free
for the callee to use, and are therefore tagged with low integrity and
low confidentiality.

 \[\mathit{Cof}(\mach,n)(\component) =
  \begin{cases}
    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
                         \text{ and for some } \funid, \codemap(\component) = \funid \\
    (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \geq \mach(\SP)
                          \text{ or if } \component = \rout\\
    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \leq \mach(\SP) \negate  n \\
    (\LOWSEC,\LOWINT) & \text{else} \\
  \end{cases}\]

Going back to our running example, recall the state of the machine at
the call from {\tt main} to {\tt g}, annotated with their
corresponding labels (not shown: the labels of the instructions
corresponding to {\tt g} that are $(\LOWSEC, \HIGHINT)$ instead of
$(\HIGHSEC, \HIGHINT)$):
\vspace*{0.2em}
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{$(\HIGHSEC,\HIGHINT)$}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\emptyoutc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{$(\HIGHSEC, \HIGHINT)$}}]%
\memory{2}{\fcallc}[{\makebox[0pt]{$(\LOWSEC, \LOWINT)$}}]%
\memory{6}{\unusedc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]
~$\cdots$
\MemoryLabel{-46.5em}{0.75em}{42}
\MemoryLabel{-42.5em}{0.75em}{0}
\MemoryLabel{-38.5em}{0.75em}{0}
\MemoryLabel{-34.5em}{0.75em}{0}
\\
\end{center}

@@

\paragraph*{Stack Integrity}

@@

\definition

Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
$\MPOTRACE$. We will say $\MPT$ satisfies {\em trace integrity} with
respect to $\contour$ if its final state (if any) agrees with its
initial one on all components marked $\HIGHSEC$ in $\contour$:
%
\[\begin{split}
  \mathit{int} ~ \contour ~ \MPT \triangleq &
    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
    \mach_e(\component) = \mach_r(\component) \\
\end{split}\]
%\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
%\sna{We could do it like this, with finiteness implicit in the existence of
%     a final state.}+1

Intuitively, the caller expects to find their private data untouched
after the callee returns. That means that if the callee never returns,
integrity is trivially satisfied. Similarly, the callee is allowed to
temporarily overwrite these locations, as long as they are restored
once control returns. We will see variations of this later on: a
stronger, inductive integrity property that ensures that the caller's
private data is never overwritten, even during the caller's execution
(a variation that leads to easier reasoning and more effective testing);
and a weaker, lazy integrity property that only ensures that the
locations that a caller accesses are preserved (a variation that allows
for a more efficient enforcement mechanism).

@@

For an example see \cref{fig:intex}, representing the unsafe substitution
\[
  44: ~ \sw ~ \rsp ~ r_4 ~ -5
\]
This instruction is in {\tt h}, and instead of storing data to the return,
it overwrites {\tt g}'s variable, {\tt z}. At the top of the figure high
integrity memory is in red, while low integrity is in green; at the end of the
trace address 103 should match the initial state, and does not. So trace
integrity is violated.
\rb{There are some color name macros, but none for integrity yet? Use instead
of hardcoded color names.}

Trace integrity characterizes the trace corresponding to a single
callee. Going back to our running example, such a trace would be, for
instance, the entire trace of {\tt h} from instruction 40 to
instruction 47. This trace is repeated in \Cref{fig:intex}, with the initial
and final states colored to show the contour; high confidentiality
components in magenta, low in teal. At the end of the trace addresses 103
and 108 have both changed; since 103 was high confidentiality, the trace
does not enjoy trace integrity.

We now lift this definition to an integrity property for a complete
execution trace. We do this by requiring that trace integrity holds
for every transition between a caller and a callee for the entirety of
the trace of the callee.

\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
a function map $\codemap$, an $n$-{\em transition} is a pair of
machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
in the trace such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point for $f$:
  $$\codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

\definition{Stack Integrity} Let $\callmap$ be a call map and
$\codemap$ a function map. We say that a system satisfies {\em stack
  integrity} with respect to $\callmap$ and $\codemap$ if, for every
initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
(if any) satisfies trace integrity with respect to the contour at the
call:
%
$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$

\paragraph*{Stack Confidentiality}

The second component of local state encapsulation is a confidentiality
property: just like integrity ensures that a caller's data is never
overwritten by a callee, confidentiality ensures that it never
influences its observable behavior. We will formally define
confidentiality in two similar steps: a trace version and a system
version.

@@

\definition{Variations}
We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
another given a contour $\contour$ if they agree on low
confidentiality data:
%
\[
\mach \approx_\contour \mach' \triangleq \forall \component .
\contour(\component) = \LOWSEC \Rightarrow \mach(\component) =
\mach'(\component)
\]
%
Then, the trace version of confidentiality operates on two traces: the
``original'' one $MP \in \MPOTRACE$ that corresponds to a callee $f$
and a ``variant'' one $M \in \MTRACE$ that would corresponding to $f$
if it started from a variant initial machine state.
%
As an example of variations, consider the top of \Cref{fig:confex1}
which shows a snapshot of the state as in the running example (top
memory layout) and a variation of it (bottom). \leo{Finalize
  contents, describe}

\definition{Trace Confidentiality} We will say two traces $MP$ and $M$
satisfy {\em trace confidentiality} with respect to $\contour$ and
a return predicate \(R \subset \MACHS\) if:
\begin{itemize}
\item If $\MPT$ is finite and its final state $R\ \mach_r$,
  then $\machT$ is also finite with some final state $\mach_r'$ and
  any changes from the initial states have changed in the same way:
$$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
  \vee \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
  \mach_r(\component) = \mach_r'(\component)$$
  In addition their observations must be equivalent:
  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
\item If $\MPT$ is infinite, then $\machT$ is also infinite and their
  observations are equivalent:
  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
\item If the $\MPT$ ends prematurely, without fulfilling \(R\), then
  we require that its observations are a prefix of those of $\machT$:
  $$\pi_o(\MPT) \lesssim \pi_o(\machT)$$
\end{itemize}

For example, consider the unsafe substitution:
\[
42 : ~ \sw ~ r_4 ~ \rsp 100
\]
Instead of reading the proper argument, {\tt h} reads directly from the
secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
in {\high} and low confidentiality in \low. Below it, the variant state
\(\mach_e'\) agrees on low confidentiality memory, but all other addresses
hold -5, in orange. After the last step, the return states \(\mach_r\) and
\(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
\(\mach_e'\). Addresses 108 and 109 have changed, shown in \tainted, and while 109
is consistent, 108 holds either 43 or -4, so the trace does not obey
trace confidentiality.
\rb{Color name macros}

Meanwhile, the following unsafe substitution outputs the secret
directly, violating the internal observation clause of confidentiality:
\[
30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
\]
%\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
%and \(\mach_e'\), below. Low confidentiality data in green is shared
%between them, while high confidentiality data in red and yellow may differ.
%After executing the code shown, the final states below have written different
%values to address 42, represented by the black and white components.
%It is allowed for address 40 to differ between the states due to the initial
%variation, but address 42 changed differently between the traces, representing
%a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
%is read, it is overwritten before return, and the write at step 3 always writes
%the same value. \Gprivc and red locations differ between variations only because
%of the initial variation, and the rest are identical. So this example passes.

@@

\definition{Stack Confidentiality}
Let $\callmap$ be a call map and $\codemap$ a function map. We say that a
system satisfies {\em stack confidentiality} with respect to
$\callmap$ and $\codemap$ if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
$f$ in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c, f, n)}
\mach_e\), let \(\machT\) be the similarly prefixed trace such that
\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, f, n)\)
and \(R = \ret{\mach_c}\):

$$\forall \MPT_{\mathit{call}} ~ \machT. ~
\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$

>>>>

In this section we formalize our definition of stack safety.We
begin with our model of \leo{compiler-provided?} knowledge about the
program structure, and then dive into the formal definitions of local state
encapsulation and well-bracketed control flow that together constitute
stack safety.

@@

\paragraph*{Return Map}
Similarly, a {\em return map} identifies addresses that complete a
return sequence. Formally, it is simply a set of words; each one
is the address of the final jump corresponding to a function return:
%
\[\retmap \in \RETMAPS \subseteq \WORDS\]
%
In our running example, the return map identifies the $\jalr$
instructions 36 and 47:
\[ \retmap = \{36, 47\} \]

@@

We separately identify, for any call state, what it means for a state
to have correctly returned from a call. We consider the call to have
returned when control returns to the instruction following
the \jal with the stack restored. Formally, this is a predicate on
states that says when the machine has just returned from a specified call state
(where $\nextPC$ gives the address of the next instruction after a given address):
\[%\begin{split}
    \ret{\mach_c} \triangleq  \lambda \mach_r . \mach_r(\PCname) =
    \nextPC(\mach_c(\PCname)) \land  \mach_r(\SP) = \mach_c(\SP)
%\end{split}
\]
There are two such return points in our running example, both of which
are depicted in the figure: the state before the execution of
instruction 27 corresponding to the return from {\tt h} and the state
before instruction 8 corresponding to the return from {\tt g}.

@@

In the rest of this section we implictly parameterize all 
definitions by a function map $\codemap$, a call map $\callmap$, a return
map $\retmap$ and an entry map $\entmap$.

@@

When \(\mach_c\) is a call state (\(\callmap(\mach_c(\PCname)) = n\) for some $n$) and
\(\mach_c \stepstoobs{\obs} \mach_e\) for some 
$\mach_e$ and $\obs$, we can construct a contour based on the bounds of the
callee's stack frame.
%
The private state of the callee is the entire area of the memory from
stack base to stack pointer, with the exception of the $n$ locations
corresponding to the arguments and result locations of the
call. Private components are marked high confidentiality and high
integrity: the callee should be able to neither read nor write them.
%
Everything above the stack pointer at the point of the call is
``uninitialized'' memory that can be written into but should not be
read. To protect the callee itself from such reads, these components
are marked high confidentiality and low integrity.
%
The same is true for the designated output location: it can be
freely written into but should not be read.
%
Special care is needed for the code memory corresponding to the
callee: they are marked low confidentiality if they belong to the
callee and high confidentiality otherwise; regardless they are always
high integrity.\apt{Sean:I think this is vestigial from an aborted version
  in which you were modeling WBCF using confidentiality. In any casen
  it is not reflected in the definition below.}
%
Finally, registers and the frame of arguments and results are free
for the callee to use, and are therefore tagged with low integrity and
low confidentiality.

 \[\mathit{Cof}(\mach,n)(\component) =
  \begin{cases}
    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
                         \text{ and } \codemap(\component) = \funid \text{ for some }\funid\\
    (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \geq \mach(\SP)
                          \text{ or if } \component = \rout\\
    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \leq \mach(\SP) \negate  n \\
    (\LOWSEC,\LOWINT) & \text{else} \\
  \end{cases}\]

Going back to our running example, here is the state of the machine at
the call from {\tt main} to {\tt g}, showing addresses annotated with their 
corresponding labels (not shown: the labels of the instructions
corresponding to {\tt g} that are $(\LOWSEC, \HIGHINT)$ instead of
$(\HIGHSEC, \HIGHINT)$)\apt{see previous note}:
\vspace*{0.2em}
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{$(\HIGHSEC,\HIGHINT)$}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\emptyoutc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{$(\HIGHSEC, \HIGHINT)$}}]%
\memory{2}{\fcallc}[{\makebox[0pt]{$(\LOWSEC, \LOWINT)$}}]%
\memory{6}{\unusedc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]
~$\cdots$
\MemoryLabel{-46.5em}{0.75em}{42}
\MemoryLabel{-42.5em}{0.75em}{0}
\MemoryLabel{-38.5em}{0.75em}{0}
\MemoryLabel{-34.5em}{0.75em}{0}
\\
\end{center}

@@

\subsubsection{Stack Integrity}

@@

\definition
Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
$\MPOTRACE$. We will say $\MPT$ satisfies {\em trace integrity} with
respect to $\contour$ if its final state (if any) agrees with its
initial one on all components marked $\HIGHSEC$ in $\contour$:
%
\[\begin{split}
  \mathit{int} ~ \contour ~ \MPT \triangleq &
    \head(\MPT) = \mpostate{\mach_e}{\_}{\_}  \land 
    \last(\MPT) = \mpostate{\mach_r}{\_}{\_} \Rightarrow \\
  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
    \mach_e(\component) = \mach_r(\component) \\
\end{split}\]
%\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
%\sna{We could do it like this, with finiteness implicit in the existence of
%     a final state.}+1

Intuitively, the caller expects to find their private data untouched
after the callee returns. That means that if the callee never returns,
integrity is trivially satisfied. Similarly, the callee is allowed to
temporarily overwrite these locations, as long as they are restored
once control returns. We will see variations of this later on: a
stronger, inductive integrity property that ensures that the caller's
private data is never overwritten, even during the caller's execution
(which leads to easier reasoning and more effective testing);
and a weaker, lazy integrity property that only ensures the preservation
of locations that subsequently affect the caller's observable actions 
(which allows for a more efficient enforcement mechanism).

@@

Trace integrity characterizes the trace corresponding to a single
callee.
Going back to our running example, such a trace would be, for
instance, the entire trace of {\tt h} from instruction 40 to
instruction 47. 
As an example of an integrity violation affecting this trace, \cref{fig:intex} shows what happens
if we change line 44 of the example program to $[\rsp - 5] \leftarrow r_4$.
Instead of storing data to the return value slot,
the changed version overwrites {\tt g}'s local variable {\tt z}.
The figure shows the modified trace, with snapshots of the initial
and final states colored to show the contour; high 
integrity memory is in red\apt{NO?}, while low integrity is in green\apt{NO?}.
At the end of the trace high integrity address 103 should match the initial state, and does not.
So trace integrity is violated.
\rb{There are some color name macros, but none for integrity yet? Use instead
  of hardcoded color names.}\apt{I don't understand what's going on with the colors.
  I thought we agreed to use the same coloring scheme throughout?}

We now lift this definition to an integrity property for a complete
execution trace. We do this by requiring that trace integrity holds
for every subtrace delimited by a call and its corresponding return.
%for every transition between a caller and a callee for the entirety of
%the trace of the callee.\apt{?????}

\apt{Some more words needed here to smooth the technicalities.}

\definition%{Transition}
Given a trace $\MPT$, an $n$-{\em transition} is a pair of
machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
in the trace such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point\apt{or is in some function?? and why do we need either??}:
  $$\exists f. \codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

\definition%{Stack Integrity}
A system satisfies {\em stack
  integrity} if, for every
initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
(if any) satisfies trace integrity with respect to the contour at the
call:
%
$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$

\subsubsection{Stack Confidentiality}

The second component of local state encapsulation is a confidentiality
property: just like integrity ensures that a caller's data is never
overwritten by a callee, confidentiality ensures that the caller's data never
influences the callee's observable behavior. We will formally define
confidentiality in two similar steps: a trace version and a system
version.

@@

\definition%{Variations}
We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
another with respect to a contour $\contour$ if they agree on low
confidentiality data:
%
\[
\mach \approx_\contour \mach' \triangleq \forall \component .
\contour(\component) = \LOWSEC \Rightarrow \mach(\component) =
\mach'(\component)
\]
%
Then, the trace version of confidentiality operates on two traces: the
``original'' tracce $MP \in \MPOTRACE$ that corresponds to a callee $f$
and a ``variant'' trace $M \in \MTRACE$ that corresponds to how $f$
would behaves if it were started from a variant initial machine state.
%
As an example of variations, consider the top of \Cref{fig:confex1}
which shows a snapshot of the state as in the running example (top
memory layout) and a variation of it (bottom). \leo{Finalize
  contents, describe}

\definition%{Trace Confidentiality}
%% Suppose $m_e \approx_\contour m_e'$.
%% Two traces $MP$ with $\head(MP) = m_e$ and $M$ with $head(M) = m_e'$
%% satisfy {\em trace confidentiality} with respect to $\contour$ and
%% a return predicate \(R \subset \MACHS\) if:
Two traces $MP$, with head $m_e$, and $M$, with head $m_e'$,
satisfy {\em trace confidentiality} with respect to contour $\contour$ and
a return predicate \(R \subset \MACHS\) if, whenever $m_e \approx_\contour m_e'$ then:
\begin{itemize}
\item If $\MPT$ is finite with final state $\mach_r$ and $\mach_r \in R$, 
  then $\machT$ is also finite with some final state $\mach_r'$ and
  any changes from the initial states have changed in the same way:
$$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
  \vee \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
  \mach_r(\component) = \mach_r'(\component)$$
  In addition their observations must be similar:
  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
\item If $\MPT$ is infinite, then $\machT$ is also infinite and their
  observations are similar:
  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
\item If the $\MPT$ ends prematurely, without fulfilling \(R\), then
  we require that its observations are a prefix of those of $\machT$:
  $$\pi_o(\MPT) \lesssim \pi_o(\machT)$$
\end{itemize}
\apt{Needs words about why we need a return predicate abstraction.}

Consider the result of modifying line 42 of our running example to $r_4 \leftarrow [\rsp - 9]$.
Instead of reading the proper argument {\tt v}, {\tt h} now reads directly from the
secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
in {\high} and low confidentiality in \low. Below it, the variant state
\(\mach_e'\) agrees on low confidentiality memory, but all other addresses
hold -5, in orange.\apt{NO?}  After the last step, the return states \(\mach_r\) and
\(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
\(\mach_e'\). Addresses 108 and 109 have changed, shown in \tainted, and while 109
had changed in a consistent way (independent of the variation), 108 holds either 43 or -4, so the trace does not obey trace confidentiality.
\rb{Color name macros}
\apt{Explain slash notation in effects column.}

As another example, 
changing line 30 of the running example to $r_5  \leftarrow [\rsp - 6]$ would cause
{\tt g} to read the value of {\tt main}'s local variable {\tt x} and then immediately
output it, violating the internal observation clause of the confidentiality definition.
%\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
%and \(\mach_e'\), below. Low confidentiality data in green is shared
%between them, while high confidentiality data in red and yellow may differ.
%After executing the code shown, the final states below have written different
%values to address 42, represented by the black and white components.
%It is allowed for address 40 to differ between the states due to the initial
%variation, but address 42 changed differently between the traces, representing
%a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
%is read, it is overwritten before return, and the write at step 3 always writes
%the same value. \Gprivc and red locations differ between variations only because
%of the initial variation, and the rest are identical. So this example passes.

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
We say that a
system satisfies {\em stack confidentiality}
% with respect to $\callmap$ and $\codemap$
if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ 
in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)}
\mach_e\), let \(\machT\) be the similarly prefixed trace such that
\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, n)\)
and \(R = \ret{\mach_c}\):

$$\forall \MPT_{\mathit{call}} ~ \machT. ~
\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, n)} \mach_e' \wedge
\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\mathit{conf}~\mathit{Cof}(\mach_c, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$

\apt{oy. still needs work to make this intelligible}
